A Large-scale Study of Polymorphism in R

### 2019-03-21 paper/main.tex
methods

@@

Our tracer is built on of the {\tt genthat}~\cite{issta18} tracer, which we
modified to capture the type information of arguments.  {\tt genthat}
generates traces for each unique function invocation with the express
purpose of re-invoking the function at a later time, and instead of
generating traces for these invocations we capture type information of the
arguments and return values of functions.  This is done by evaluating the
arguments and return values of functions at the end of the function's scope
in the relevant environment.  The sort of types and granularity of the type
information that are generated can be specified in our tracer: Essentially,
our tracer is parameterized over a user-specified type system.  For
instance, if desired our tracer can dig into lists and data frames to get
the types of list elements and fields.  Our tracer produces trace results
for each unique type signature encountered, which may then be used for data
analysis.

>>>>

\newcommand{\genthat}{{\sc genthat}\xspace}

@@

The basic idea we are exploring is to observe arguments and return values of
function calls, and from these generalize possible type signatures for the
called functions. To this end, we build on an open source tool called
\genthat whose purpose was to generate unit tests for R
libraries~\cite{issta18}.  To generate test cases, one must record and
recreate accurately arguments and return values and avoid generating
redundant tests (i.e. tests that do not increase code coverage). For our
purposes, we change the existing tool in two main ways, we record argument
\emph{shapes} rather than their values, given the limited number of shapes
this allowed us to avoid having to run code coverage minimization phase of
the tool. Both changes were beneficial for scalability -- allowing us to
trace more calls than was reported in~\cite{issta18}.

\subsection{Example}

To illustrate our approach consider the following script which adds a double
to an integer, and the creates a matrix and a vector to i.

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
> 1 + 2L
[1] 3
> x <- c(1,2,3,4)
> attr(x,"dim") <- c(2,2)
> x + c(1,2)
     [,1] [,2]
[1,]    2    4
[2,]    4    6
\end{lstlisting}}\caption{Example script}\label{example}\end{figure}

The functions being called here are \k{`+`}, \k{`<-`}, \k{c} and
\k{`attr<-`}.  Some functions syntactic support, addition is infix, and the
assignment to an attribute is written mixfix. The shapes we would expect to
record are as follows:

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
`+`: D[1], I[1] -> D[1]
`c`: D[1], D[1], D[1], D[1] -> D[4]
`<-`: S, D[3] -> D[3]
`c`: D[1], D[1] -> D[2]
`attr<-1`:  S, C, D[2] -> 
`c`: D[1], D[1] -> D
`+`: D[4]{dim=(2,2)}, D[2] -> D[4]{dim=(2,2)}
\end{lstlisting}}\caption{Recorded shapes}\label{shapes}\end{figure}

In the above we abreviate types, \k{double}, \k{integer}, \k{symbol} and
\k{character}. For vectorized types we record their length and for all types
we record attributes with some of their values. Looking at the signatures
observed for addition, it is clear that the function is polymorphic as it
starts with the addition of two scalar numbers of different types, and then
adds a matrix to a vector returning a matrix.

\subsection{Implementation}

tracer, which we
modified to capture the type information of arguments.  {\tt genthat}
generates traces for each unique function invocation with the express
purpose of re-invoking the function at a later time, and instead of
generating traces for these invocations we capture type information of the
arguments and return values of functions.  This is done by evaluating the
arguments and return values of functions at the end of the function's scope
in the relevant environment.  The sort of types and granularity of the type
information that are generated can be specified in our tracer: Essentially,
our tracer is parameterized over a user-specified type system.  For
instance, if desired our tracer can dig into lists and data frames to get
the types of list elements and fields.  Our tracer produces trace results
for each unique type signature encountered, which may then be used for data
analysis.
